Forecasting League of Legends Player Performance through In-Game Metrics and Key Statistics

### 2024-11-26 paper/paper.qmd
update the paper.qmd, model part and train test validation part

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(arrow)
```

@@

@@

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned league of legend esport data
#| echo: false
#| warning: false
#| tbl-width: 15
#| tblg-height: 8

@@

```{r}

@@

summary(model_MLR)

@@

```{r}
# Multiple R-squared:  0.337
```

## Bayesian Model set-up {#sec-modelsetup}

The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. We run the model with the following specifications:

- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$

- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$

- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(arrow)
library(rstanarm)
library(ggplot2)
```

@@

The data simulation is done in python 3.9[@citePython] with the following packages: Numpy [@citeNumPy] and pandas[@citepandas].

@@

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned league of legend esport data
#| echo: false
#| warning: false
#| tbl-width: 15
#| tbl-height: 8

@@

```{r}
#| label: fig-model-MLR
#| fig-cap: "The MLR shows better performance on captured significant variance in players KDA performance."
#| echo: false
#| warning: false

@@

# summary(model_MLR)

@@

The plot @fig-model-MLR visualizes the results of a Multiple Linear Regression (MLR) model predicting KDA using predictors such as side, vision score, total gold, total CS, damage to champions, and game duration. The x-axis represents the actual KDA values, while the y-axis represents the predicted KDA values. Each dot corresponds to an individual game, with colors indicating different players. The red dashed line represents the perfect prediction line (y=x), where predicted values would perfectly match the actual values.

The model demonstrates an improved ability to capture variance in KDA compared to a simple linear regression model, as evidenced by many points clustering closer to the red dashed line.

### Compute R Squared for SLR Model
\begin{align*}
SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
\text{R-squared} = 0.337
\end{align*}

We could see that there is a strong improvement on the R-squared values, which mean the MLR model have better performance on illustration for the response variable but still reflects limited predictive accuracy, as only 33.7% of the variance in the KDA is explained by the MLR model.

## Bayesian Model

### Bayesian Model Set-UP
\begin{align*}
\text{KDA}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_1 \cdot \text{visionscore}_i + \beta_2 \cdot \text{totalgold}_i + \beta_3 \cdot \text{totalcs}_i \\
&+ \beta_4 \cdot \text{damagetochampions}_i\\
&+ u_{\text{side}[i]} + u_{\text{gamelength}[i]} + u_{\text{position}[i]} \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta_j &\sim \text{Normal}(0, 2.5), \quad j = 1, 2, 3, 4 \\
u_{\text{side}} &\sim \text{Normal}(0, \sigma_{\text{side}}) \\
u_{\text{gamelength}} &\sim \text{Normal}(0, \sigma_{\text{gamelength}}) \\
u_{\text{position}} &\sim \text{Normal}(0, \sigma_{\text{position}})
\end{align*}

This hierarchical Bayesian model is designed to capture both the fixed effects of predictors and the random effects of grouping factors. The response variable is KDA, predictors are visionscore, totalgold, total.cs, damagetochampions, side,gamelength and position. A normal prior with mean 0 and standard deviation 2.5 is set for all coefficients and the intercept, scaled automatically.

### Posterior Predictive Checks

```{r}
#| echo: false
#| warning: false
# formula = KDA ~ visionscore + totalgold + total.cs + damagetochampions + gamelength + (1 | side)
set.seed(123)

bayesian_model_1 <- readRDS(here("models/bayesian_model_1.rds"))

# pp_check(bayesian_model_1)
# bayesian_model_predictions = posterior_predict(bayesian_model_1, newdata = cleaned_lol_data)
# ss_total <- sum((cleaned_lol_data$KDA - mean(cleaned_lol_data$KDA))^2)
# ss_residual <- sum((cleaned_lol_data$KDA - colMeans(bayesian_model_predictions))^2)
# r_squared <- 1 - (ss_residual / ss_total)
# print(r_squared)
```

```{r}
#| echo: false
#| warning: false

# # Add predictions to the data
# cleaned_lol_data$predicted_KDA <- apply(bayesian_model_predictions, 2, mean) # Take the mean of posterior predictions

# # Create a visualization with actual vs. predicted values
# ggplot(cleaned_lol_data, aes(x = damagetochampions)) +
#   geom_line(aes(y = KDA, color = "Actual"), size = 1) + # Line for actual values
#   geom_line(aes(y = predicted_KDA, color = "Predicted"), size = 1, linetype = "dashed") + # Line for predicted values
#   labs(
#     title = "Actual vs Predicted KDA",
#     x = "Game Index",
#     y = "KDA",
#     color = "Legend"
#   ) +
#   theme_minimal() +
#   theme(
#     legend.position = "bottom"
#   )

```{r}
#| label: fig-bayesian-model
#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
#| echo: false
#| warning: false
pp_check(bayesian_model_1)
```

@fig-bayesian-model The posterior predictive check (PPC) plot compares the observed KDA distribution $y$(dark blue line) with the posterior predictive distributions $y_{rep}$(lighter blue lines) generated by the Bayesian model. The alignment between the observed and predictive distributions suggests that the model captures the central tendencies and overall variability of KDA effectively. The peak and general shape of the observed distribution are closely mirrored by the simulated predictions, indicating that the included predictors—vision score, total gold, total CS, damage to champions, position and game length are relevant in explaining the main patterns in KDA.

However, the model shows slight underperformance in capturing extreme values, as evidenced by discrepancies in the tails of the distributions.

In conclusion, the Bayesian model performs well in approximating the central tendencies and variance of KDA, as demonstrated by the close alignment between observed and predictive distributions. However, some refinements may be necessary to improve the model's ability to capture extreme values and better represent the full range of variability in the data.

```{r}
#| echo: false
#| warning: false
# pp_check(bayesian_model_1)

# summary(bayesian_model_1)
# bayesian_model_predictions = posterior_predict(bayesian_model_1, newdata = cleaned_lol_data)
```

```{r}
#| warning: false
#| echo: false
# ss_total <- sum((cleaned_lol_data$KDA - mean(cleaned_lol_data$KDA))^2)
# ss_residual <- sum((cleaned_lol_data$KDA - colMeans(bayesian_model_predictions))^2)
# r_squared <- 1 - (ss_residual / ss_total)
```

### Train Test Validation
we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting KDA for BLG players. The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively. We used the stan_glmer function to fit the model on the training data, with predictors includingvision score, total gold, total CS, damage to champions, position and game length. After training, we test the model by generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy.

```{r}
#| echo: false
#| warning: false
train_data = read_parquet(here("data/02-analysis_data/train_data.parquet"))
test_data = read_parquet(here("data/02-analysis_data/test_data.parquet"))

bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))

predictions <- posterior_predict(bayesian_model_train, newdata = test_data)

ss_total <- sum((test_data$KDA - mean(test_data$KDA))^2)
ss_residual <- sum((test_data$KDA - colMeans(predictions))^2)
r_squared <- 1 - (ss_residual / ss_total)

# R-Squared: 0.6281344
```

```{r}
print(r_squared)
```

\begin{align}
SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
\text{R-squared} = 0.700563202695799
\end{align}

Where:\
\begin{itemize}
    \item \( y_i \) represents the actual values,
    \item \( \bar{y} \) is the mean of the actual values,
    \item \( \hat{y}_i \) are the predicted values,
    \item \( SS_{\text{total}} \) is the total sum of squares,
    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
\end{itemize}
